Fast-Match:\\Fast and Robust Feature Matching on Large Images

### 2014-06-20 paper.tex
Added submission number
Made various small improvements

% TODO: Insert submission number
\def\ACCV14SubNumber{***}  % Insert your submission number here

@@

The problem that \emph{Fast-Match} attempts to solve is two-fold. By matching only image areas that are likely to correspond we hope to improve accuracy by entirely ignoring parts of the images that would otherwise likely be a source of incorrect correspondences. At the same time, this enables us to improve computational speed by not having to compute keypoints and descriptors for large parts of the images and at the same time reducing the amount of feature points we need to match. Both problems have been addressed in part by past work.

@@

Efforts to reduce the computational costs of finding nearest neighbours to feature points have largely focused on metric trees. Na\"ively the set of nearest neighbours between features in two images can computed by brute force in $O(n^2)$, where $n$ is the total number of feature points in the two images (a typical three megapixel image may contain anywhere from 500 to 5000 feature points). When SIFT was originally published Lowe, proposed using the Best-Bin-First method to approximately search for nearest neighbours \cite{beis1997shape,lowe1999object}. This reduces the computational complexity to $O(n\log n)$, but even approximate metric trees are hard pressed to compete with brute force due to the high dimensionality of SIFT and the constant costs incurred with constructing and searching in a metric tree. Later work by Muja and Lowe \cite{muja2009fast} focused on improving approximate nearest neighbour searches by using several KD-Trees simultaneously while optimizing the tree structure using K-Means to cluster similar features. Recent work on knn-graphs shows a lot of promise for high dimensional cases \cite{dong2011efficient}. We later review these improvements and their effect on efficiently matching large scale images.

@@

For sparse local image features many solutions have combined \emph{Ratio-Match} with various geometric constraints to improve matching. These constraints are based on assumptions regarding the transformation between the \emph{query} and \emph{target images}. A commonly used example is \emph{RANSAC} applied to feature matching where matches are chosen from a pool of candidates according to how well they approximate a global epipolar geometry \cite{fischler1981ransac,torr2000mlesac,szeliski2010}. Similar global angular and distance constraints can be used to filter a set of matches as shown in \cite{kim2008efficient,schmid1997local}. Finally the problem of feature matching can be modelled as a graph matching problem where each feature is a vertex, and edge values correspond to a geometric relation between two features as demonstrated in \cite{torresani2008feature,yarkony2010covering,cho2010reweighted}. For cases that require more flexibility such as scenes with moving elements and non-planar perspective change Pairwise constraints provide an alternative to the more rigid global constraints. Introduced by Leordeanu and Herbert \cite{leordeanu2005spectral}, pairwise constraints work by applying a geometric constraint across matches on a pairwise basis, attempting to minimize the pairwise error. This approach has later been adopted by Pang et al.\ but made scale invariant and more efficient by using a voting scheme \cite{yuan2012efficient,pang2012scale}. Similarly we can cluster matches together based on a geometric constraint and treat each cluster as a separate case \cite{cho2009feature,wu2011robust}. An interesting application of this principle is demonstrated by Chen et al.\ who iteratively use a Hough transform to cluster matches using angular constraints \cite{chen2013robust}.

@@

\subsection{Considerations}
\label{considerations}
%
If we set out to design a truly fast matching algorithm, we cannot rely just on optimizing the matching step once the descriptors from both images have been found and computed. Finding and computing descriptors can easily account for 80\% of the time spent matching for bigger images (cf.\ Figure~\ref{fig:timings}). For this reason \emph{Fast-Match} is designed to only compute features for the part of the image we hope to match.

@@

We a \emph{general} and a \emph{retrieval} variation of the algorithm. The latter assumes that we know one of the images that we intend to match ahead of time and can do some off-line computations. For the \emph{general} variation we make those computations on the fly instead and do not assume anything to be pre-computed.

\begin{algorithm}[htb]
\caption{Fast-Match}
\label{alg-fast}
\begin{algorithmic}
    \Require $I_{query}, I_{target}$ : images, maxiter $\in \mathbb{N}$, $\tau \in [0,1]$
    \State $M_{seed} ~\gets$ seed\_matches$(I_{query}, I_{target})$
    \State $M_{final} \gets \varnothing$
    \State $C_{final}~ \gets \varnothing$
    \State $M_{seen} ~\gets \varnothing$
    \While{$M_{seed} \ne \varnothing \wedge i < $maxiter$}$
        \State $M_{round}~ \gets$ get\_matches$(M_{seed})$
        \State $C_{round}~\,\gets$ get\_confidence$(M_{round})$
        \State $M_{seed}~~~ \gets$ get\_seeds$(M_{round} \setminus M_{seen}, C_{round}, \tau)$
        \State $M_{seen}~~\,\gets M_{seen} \cup M_{seed}$
        \State $M_{final}~\,\gets M_{final} \cup M_{round}$
        \State $C_{final}~~\,\gets C_{final} \cup C_{round}$
    \EndWhile \\
    \Return $M_{final}, C_{final}$
\end{algorithmic}
\end{algorithm}

\subsection{Initiating Seeds}
\label{seeds}
%
Depending on the scenario, several strategies can be used to obtain a set of initial seed matches. For the case of matching a pair of two images we can get a rough group of seed matches by matching the thumbnails of the two images with for example \emph{Ratio-Match}. However if we were matching images in a series, such as frames from a movie we could instead make use of a subset of the matches from the last frame to seed \emph{Fast-Match} on the next. In practice we have found it efficient to resize both images to thumbnails of $300\times 300$ pixels and use ratio match to obtain a set of matches and ratios that we then threshold with the confidence threshold $\tau$ to obtain the initial seed matches.

@@

Lowe and others have shown that the distance between two SIFT descriptors is much less indicative of a true correspondence than the ratio between the best and second best match \cite{lowe2004sift,mikolajczyk2005performance,moreels2007evaluation,rabin2009statistical}. This ratio is more formally defined as follows: If we let $f_q$ be a feature in the \emph{query image} and $f_t, f_b$ be the two nearest neighbours of $f_q$ in the \emph{target image} then the ratio $r$ is defined as follows:
\begin{align*}
    r &= %\text{r}(f_{q}, (f_t, f_b)) \\
        &= \frac{d(f_{q}, f_{t})}{d(f_{q}, f_{b})}.
\end{align*}
Here $d(f_i,f_j)$ is the distance between the features $f_i$ and $f_j$. For SIFT this is the Euclidean distance. The distance between two feature points is greater when the feature points resemble each other less. For this reason we have higher confidence in a match with a low $r$-value. Using $r$ as a measure of match confidence presumes that we expect each feature in the \emph{target image} to have at most one true correspondence in the \emph{target image}. Intuitively if we try to find a match for a feature $f_i$ in an image that does not have any true correspondences, then we would expect the two closest neighbours to be roughly equally well matched with $f_i$. For this reason we attribute high confidence to matches where the closest neighbour is dramatically closer to $f_i$ than the second closest neighbour and discard the rest.

@@

We will show that the number of possible seed matches is on average linear in the number of image features, and we also provide an upper bound for the probability that it is not. We assume that outside of true correspondences, features have better or at least equally good matches in terms of confidence in the image they come from. More rigorously put: For any feature in the target image $f_t$ let $f_1, f_2, \ldots$ be the best, second best, etc matching feature from either image which is not a true correspondence. Let $A_{ti}$ be the event that $f_i$ is found in the \emph{query image}, then $\mathbb{P}(A_{ti}) \leq 0.5$ and $A_{ti}$ is independent from $A_{tj}$ when $i \ne j$.

@@

In the worst case, each feature in the \emph{target image} has a true correspondence in the \emph{query image} and $\mathbb{P}(A_{ti} = 0.5)$, in which case we find can find $\mathbb{P}(X_t = n) = \mathbb{P}(A_{ti})^n = 0.5^n$, as well as the expected value $\mathbb{E}(X_t) = \sum_{i=1}^\infty i0.5^i = 2$, and the variance $Var(X_t) = \sum_{i=1}^\infty 0.5^i(i - 2)^2 = 6$.

@@

Figure~\ref{fig:timings} demonstrate the speed of \emph{Fast-Match} on image sizes as is typically produced by modern consumer cameras and camera phones. In Figure~\ref{fig:timings} we compare the speed of the two variations of \emph{Fast-Match} to different variations of \emph{Ratio-Match} on two image pairs, one with two images of 10 megapixel and another where the same image pair has been scaled to 3 megapixels. For the larger images we see the clearest difference in performance. The retrieval variant of \emph{Fast-Match} matches the image pairs in around one second while a retrieval variant of \emph{Ratio-Match} with precomputed features spends eight seconds on the same task. This result is roughly equal to the time the general variation of \emph{Fast-Match} spends matching the two images without any pre-computations. For the nearest neighbor search we note that for large images there are substantial speed gains to be had by choosing the right algorithm for finding nearest neighbors, with Muja et al.'s Flann matcher being vastly superior to brute force and KGraph.

@@

Figure~\ref{fig:match_example} shows the result from matching the pair of three megapixel images using \emph{Fast-Match} and \emph{Ratio-Match}. %An almost identical result can be obtained using the pair of 10 megapixel pictures instead. 
\emph{Fast-Match} ends up only matching the part of the images that are likely to have correspondences. This highlights both a strength and a weakness of \emph{Fast-Match}. It is this myopic matching approach that is responsible for \emph{Fast-Match}' speed and precision; on the other hand, \emph{Fast-Match} is not the best candidate for images that are almost identical, because its step by step approach incurs too much overhead to compete in speed with \emph{Ratio-Match} in those cases.

>>>>

% TODO: Insert submission number
\def\ACCV14SubNumber{325}  % Insert your submission number here

@@

The problem that \emph{Fast-Match} attempts to solve is two-fold. By matching only image areas that are likely to correspond we hope to improve accuracy by entirely ignoring parts of the images that would otherwise likely be a source of incorrect correspondences. At the same time, this enables us to improve computational speed by not having to compute keypoints and descriptors for large parts of the images and at the same time reducing the amount of feature points we need to match.

@@

Efforts to reduce the computational costs of finding nearest neighbours
to feature points have largely focused on metric trees. Na\"ively the
set of nearest neighbours between features in two images can computed by
brute force in $O(n^2)$, where $n$ is the total number of feature points
in the two images (a typical three megapixel image may contain anywhere
from 500 to 5000 feature points). When SIFT was originally published,
Lowe proposed using the Best-Bin-First method to approximately search
for nearest neighbours \cite{beis1997shape,lowe1999object}. This reduces
the computational complexity to $O(n\log n)$, but for smaller images
even approximate metric trees are hard pressed to compete with brute
force due to the high dimensionality of SIFT and the costs
incurred with constructing the metric tree. Later work by Muja and Lowe \cite{muja2009fast} focused on improving approximate nearest neighbour searches by using several KD-Trees simultaneously while optimizing the tree structure using K-Means to cluster similar features. Recent work on knn-graphs shows a lot of promise for high dimensional cases \cite{dong2011efficient}. We later review these improvements and their effect on efficiently matching large scale images.

@@

For sparse local image features many solutions have combined \emph{Ratio-Match} with various geometric constraints to improve matching. These constraints are based on assumptions regarding the transformation between the \emph{query} and \emph{target images}. A commonly used example is \emph{RANSAC} applied to feature matching where matches are chosen from a pool of candidates according to how well they approximate a global epipolar geometry \cite{fischler1981ransac,torr2000mlesac,szeliski2010}. Similar global angular and distance constraints can be used to filter a set of matches as shown in \cite{kim2008efficient,schmid1997local}. Finally the problem of feature matching can be modelled as a graph matching problem where each feature is a vertex, and edge values correspond to a geometric relation between two features as demonstrated in \cite{torresani2008feature,yarkony2010covering,cho2010reweighted}. For cases that require more flexibility such as scenes with moving elements and non-planar perspective change pairwise constraints provide an alternative to the more rigid global constraints. Introduced by Leordeanu and Herbert \cite{leordeanu2005spectral}, pairwise constraints work by applying a geometric constraint across matches on a pairwise basis, attempting to minimize the pairwise error. This approach has later been adopted by Pang et al.\ but made scale invariant and more efficient by using a voting scheme \cite{yuan2012efficient,pang2012scale}. Similarly we can cluster matches together based on a geometric constraint and treat each cluster as a separate case \cite{cho2009feature,wu2011robust}. An interesting application of this principle is demonstrated by Chen et al.\ who iteratively use a Hough transform to cluster matches using angular constraints \cite{chen2013robust}.

@@

\subsection{Considerations}
\label{considerations}
%
If we set out to design a truly fast matching algorithm, we cannot rely
just on optimizing the matching step once the descriptors from both
images have been found and computed. Finding and computing descriptors
can easily account for 80\% of the time spent matching in the case of bigger images (cf.\ Figure~\ref{fig:timings}). For this reason \emph{Fast-Match} is designed to only compute features for the part of the image we hope to match.

@@

We introduce a \emph{general} and a \emph{retrieval} variation of the algorithm. The latter assumes that we know one of the images that we intend to match ahead of time and can do some off-line computations. For the \emph{general} variation we make those computations on the fly instead and do not assume anything to be pre-computed.

\begin{algorithm}[htb]
\caption{Fast-Match}
\label{alg-fast}
\begin{algorithmic}
    \Require $I_{query}, I_{target}$ : images, maxiter $\in \mathbb{N}$, $\tau \in [0,1]$
    \State $M_{seed} ~\gets$ seed\_matches$(I_{query}, I_{target})$
    \State $M_{final} \gets \varnothing$
    \State $C_{final}~ \gets \varnothing$
    \State $M_{seen} ~\gets \varnothing$
    \While{$M_{seed} \ne \varnothing \wedge i < $maxiter$}$
        \State $M_{round}~ \gets$ get\_matches$(M_{seed})$
        \State $C_{round}~\,\gets$ get\_confidence$(M_{round})$
        \State $M_{seed}~~~ \gets$ get\_seeds$(M_{round} \setminus M_{seen}, C_{round}, \tau)$
        \State $M_{seen}~~\,\gets M_{seen} \cup M_{seed}$
        \State $M_{final}~\,\gets M_{final} \cup M_{round}$
        \State $C_{final}~~\,\gets C_{final} \cup C_{round}$
    \EndWhile \\
    \Return $M_{final}, C_{final}$
\end{algorithmic}
\end{algorithm}
%
\subsection{Initiating Seeds}
\label{seeds}
%
Depending on the scenario, several strategies can be used to obtain a set of initial seed matches. For the case of matching a pair of two images we can get a rough group of seed matches by matching the thumbnails of the two images with for example \emph{Ratio-Match}. However if we were matching images in a series, such as frames from a movie we could instead make use of a subset of the matches from the last frame to seed \emph{Fast-Match} on the next. In practice we have found it efficient to resize both images to thumbnails of $300\times 300$ pixels and use ratio match to obtain a set of matches and ratios that we then threshold with the confidence threshold $\tau$ to obtain the initial seed matches.

@@

Lowe and others have shown that the distance between two SIFT descriptors is much less indicative of a true correspondence than the ratio between the best and second best match \cite{lowe2004sift,mikolajczyk2005performance,moreels2007evaluation,rabin2009statistical}. This ratio is more formally defined as follows: If we let $f_q$ be a feature in the \emph{query image} and $f_t, f_b$ be the two nearest neighbours of $f_q$ in the \emph{target image} then the ratio $r$ is defined as follows:
\begin{align*}
    r &= \frac{d(f_{q}, f_{t})}{d(f_{q}, f_{b})}.
\end{align*}
Here $d(f_i,f_j)$ is the distance between the features $f_i$ and $f_j$. For SIFT this is the Euclidean distance. The distance between two feature points is greater when the feature points resemble each other less. For this reason we have higher confidence in a match with a low $r$-value. Using $r$ as a measure of match confidence presumes that we expect each feature in the \emph{target image} to have at most one true correspondence in the \emph{target image}. Intuitively if we try to find a match for a feature $f_i$ in an image that does not have any true correspondences, then we would expect the two closest neighbours to be roughly equally well matched with $f_i$. For this reason we attribute high confidence to matches where the closest neighbour is dramatically closer to $f_i$ than the second closest neighbour and discard the rest.

@@

For the overall running time to be linear, we will need to show that the
number of possible seed matches is linear in the number of image
features. We will prove that this is the case on average and provide an
upper bound for the probability that the average case will not happen. We assume that outside of true correspondences, features have better or at least equally good matches in terms of confidence in the image they come from. More rigorously put: For any feature in the target image $f_t$ let $f_1, f_2, \ldots$ be the best, second best, etc matching feature from either image which is not a true correspondence. Let $A_{ti}$ be the event that $f_i$ is found in the \emph{query image}, then $\mathbb{P}(A_{ti}) \leq 0.5$ and $A_{ti}$ is independent from $A_{tj}$ when $i \ne j$.

@@

In the worst case, each feature in the \emph{target image} has a true
correspondence in the \emph{query image} and $\mathbb{P}(A_{ti} = 0.5)$,
in which case we compute the probability that $X_t = n$ as well as the
expected value and variance of $X_t$:
\begin{align}
    \mathbb{P}(X_t = n) &= \mathbb{P}(A_{t1} \wedge \ldots \wedge A_{tn}) = 0.5^n \\
    \mathbb{E}(X_t) &= \sum_{i=1}^\infty i0.5^i = 2 \\
    Var(X_t) &= \sum_{i=1}^\infty 0.5^i(i - 2)^2 = 6
\end{align}

@@

Figure~\ref{fig:timings} demonstrate the speed of \emph{Fast-Match} on
image sizes as is typically produced by modern consumer cameras and
camera phones. We compare the speed of the two variations of \emph{Fast-Match} to different variations of \emph{Ratio-Match} on two image pairs, one with two images of 10 megapixel and another where the same image pair has been scaled to 3 megapixels. For the larger images we see the clearest difference in performance. The retrieval variant of \emph{Fast-Match} matches the image pairs in around one second while a retrieval variant of \emph{Ratio-Match} with precomputed features spends eight seconds on the same task. This result is roughly equal to the time the general variation of \emph{Fast-Match} spends matching the two images without any pre-computations. For the nearest neighbor search we note that for large images there are substantial speed gains to be had by choosing the right algorithm for finding nearest neighbors, with Muja et al.'s Flann matcher being vastly superior to brute force and KGraph.

@@

Figure~\ref{fig:match_example} shows the result from matching the pair of three megapixel images using \emph{Fast-Match} and \emph{Ratio-Match}. %An almost identical result can be obtained using the pair of 10 megapixel pictures instead. 
\emph{Fast-Match} ends up only matching the part of the images that are
likely to have correspondences. This highlights both a strength and a
weakness of \emph{Fast-Match}. It is the narrow matching approach which is responsible for \emph{Fast-Match}' speed and precision; on the other hand, \emph{Fast-Match} is not the best candidate for images that are almost identical, because its step by step approach incurs too much overhead to compete in speed with \emph{Ratio-Match} in those cases.
